---
title: "Lake Water Quality Analysis"
author: "Mikaela Richardson"
date: "2/24/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(lubridate) #For dealing with date and time
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
#lagosne_get(dest_folder = LAGOSNE:::lagos_path(),overwrite=T)

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
#names(nutr)
```

### Subset columns nutr to only keep key info that we want


```{r}
clarity_only <- nutr %>%
  select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% ymd(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observatiosn did we lose?
# nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)


```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                            distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')
```

### Mean Chl_a map

```{r}
### Take the mean chl_a and secchi by lake

mean_values_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,mean_values_200,
                          by='lagoslakeid') 

#Make a map
mapview(mean_spatial,zcol='log10_mean_chl')
```


# Class work

## 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for sites with at least 200 observations?

- Here, I just want a plot of chla vs secchi for all sites 

```{r, warning = F}
#Sites with at least 200 observations
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200) 

cor(chla_secchi_200$chla, chla_secchi_200$secchi)

ggplot(chla_secchi_200, aes(x=chla, y= secchi))+
  geom_point()+
  labs(x = "log(Chlorophyll A)", y = "log(Secchi Disk Depth)", title = "Chorophyll A vs. Secchi Disk Depth")+
  scale_x_log10()+
  scale_y_log10()
```

Figure 1. Scatterplot of Chlorophyll A and Secchi disk depth for all sites. The log scale was taken for both variables to have a better understanding of the correlation between the two. There is a negative correlation between chlorophyll a and the secchi disk depth with a correlation value of -0.3774. As the visible depth of the secchi disk increases there is less chlorophyll a. With more chlorophyll a the water will be more turbid, so it would be expected that the secchi disk depth would decrease as it would more difficult to see.

## 2) What states have the most data? 

Minnesota has the most data collected. 

### 2a) First you will need to make a lagos spatial dataset that has the total number of counts per site.

```{r}
site_counts <- nutr %>%
  group_by(lagoslakeid)%>%
  summarize(count = n())

spatial_counts <- inner_join(spatial_lakes, site_counts %>%
                               distinct(lagoslakeid,.keep_all=T),
                             by = "lagoslakeid")
```


### 2b) Second, you will need to join this point dataset to the us_boundaries data. 

```{r}
states_lagos <- lagos$state

states_lakes <- inner_join(spatial_counts, states_lagos, by = "state_zoneid") 
```


### 2c) Then you will want to group by state and sum all the observations in that state and arrange that data from most to least total observations per state. 

```{r}
sum_states <- states_lakes %>% 
  as.data.frame()%>%
  select(-geometry)%>%
  group_by(state_name) %>%
  summarize(sum_counts = sum(count))%>%
  arrange(desc(sum_counts))

head(sum_states)
```
Accessed code for lines 164, 174, and 175 from the Code Help page on the class' Teams group by Matt Ross and Amanda Hastings on 2/24/2022 and for lines 155-157 from Matt Ross' Day 8 video lecture.

## 3) Is there a spatial pattern in Secchi disk depth for lakes with at least 200 observations?

There isn't a clear spatial pattern for mean Secchi disk depth for lakes with at least 200 observations, however, there does seem to be lower values for the Midwest and south of Boston. With more farming in the Midwest, it would be expected that there is more nutrient loading in the lakes increasing chlorophyll a concentrations and reducing the Secchi disk depth.

```{r}
mean_values_200 <- chla_secchi_200 %>%
  group_by(lagoslakeid) %>%
  summarize(mean_secchi = mean(secchi, na.rm = T),
            mean_chl = mean(chla, na.rm =T)) %>%
  mutate(log10_mean_secchi = log10(mean_secchi))

nutr_spatial <- inner_join(spatial_lakes, mean_values_200 %>%
                             distinct(lagoslakeid, .keep_all=T),
                           by = "lagoslakeid")
 
 mapview(nutr_spatial,zcol = 'log10_mean_secchi')

```


